Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-08-09 ms.tex
steve warrens edits and forgot figures long time ago

Here, we use the \citet{Wogan_2023} results, along with Monte-Carlo simulations of Earth's impact history to make an alternative estimate for when life most likely emerged during an RNA-first process. Our calculations account for the possibility of planet sterilization by impacts that vaporize the ocean \citep{Sleep_1989}. We assume that a life-starting impact is one that produces significant prebiotic nitriles and is not subsequently followed by an ocean-vaporizing impact that destroys the biosphere. By considering the fraction of stochastic impact realizations that do not have an life-starting impact, we also estimate the probability of life beginning if Earth history was rerun.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and log-linearly extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

@@

\begin{equation}
  N_0 = A_\oplus s_0 n_m = A_\oplus s_0 10^{a e^{b t} + c}
  \label{eq:num_earth_impacts}
\end{equation}
Equation \eqref{eq:num_earth_impacts} also accounts for Earth's surface area ($A_\oplus$), making $N_0$ the number of impacts on Earth since age $t$ that cause a crater $> 1$ km on the moon. As discussed previously, \citet{Morbidelli_2018} use crater scaling relations to show that a 1 km lunar crater corresponds to a 50 m object with mass $m_0 = 1.64 \times 10^8$ kg.

@@

\begin{equation}
  \overline{N}_{ij} = N(m_{j-\frac{1}{2}},t_{i-\frac{1}{2}},t_{i+\frac{1}{2}}) - N(m_{j+\frac{1}{2}},t_{i-\frac{1}{2}},t_{i+\frac{1}{2}})
\end{equation}
Next, we sample a Poisson distribution for each $\overline{N}_{ij}$, giving a stochastic number of impacts in each mass and time grid cell which constitutes an impact history. By modeling impacts as a Poisson process, we are assuming that each impact does not change the probability of the next one (i.e., impacts are independent). Performing this sampling thousands of times captures many of the possible impact histories. We only keep impact histories that accrete a total mass of $2 \times 10^{22}$ and $6 \times 10^{22}$ kg because this is approximately the mass implied by the HSEs in Earth's mantle, assuming that HSEs were delivered by late accretion. This range of masses is based on \citet{Day_2015} who used mantle HSEs to suggest that the Earth was impacted by $\sim 3 \times 10^{22}$ to $4.8 \times 10^{22}$ kg of material, but we choose a wider range of masses because HSEs could have been lost to Earth's core or to space during impacts \citep{Marchi_2018}.

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Appendix Figure \ref{fig:velocity_distribution} shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10 km$). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would cause biopoiesis depends on the required minimum impact mass to produce significant origin of life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} uses photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} finds that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.

@@

As illustrated in Figure \ref{fig:example_impact_histories}, a life-starting impact does not occur in every simulated impact history. In our impact-driven model for the origin of life, biopoiesis requires (1) an impact of sufficient mass to make prebiotic nitriles and (2) the lack of a later ocean-vaporizing collision that destroys the biosphere without rekindling it. Figure \ref{fig:probabilities_of_impacts} shows the probability of both conditions occurring as a function of the minimum impact mass that produces prebiotic molecules. For example, consider the \citet{Wogan_2023} optimistic minimum mass of $4 \times 10^{20}$ kg indicated in Figure \ref{fig:probabilities_of_impacts}. All simulated impact histories have at least one impact bigger than $4 \times 10^{20}$ kg (blue line in Figure \ref{fig:probabilities_of_impacts} is 1.0 at the given mass). However, the probability of the origin of life in this scenario is 92\%, because 8\% of the time the last $4 \times 10^{20}$ kg collision is followed by a smaller ocean-vaporizing impact that perhaps sterilizes the planet (orange dashed line in Figure \ref{fig:probabilities_of_impacts} with 0.92 at $4 \times 10^{20}$ kg). An origin of life is less probable if we consider the \citet{Wogan_2023} pessimistic minimum mass ($> 5 \times 10^{21}$ kg) to produce significant origin of life nitriles. A $> 5 \times 10^{21}$ kg impact without later ocean vaporization occurs 41\% of the time. A life-starting impact was therefore relatively likely on the early Earth, and was strongly favored (92\% probability) under optimistic assumptions.

@@

To estimate the most likely timing for the origin of life, we consider both the \citet{Wogan_2023} optimistic and pessimistic criteria. Figure \ref{fig:time_and_mass}a is the optimistic case, showing the timing (Figure \ref{fig:time_and_mass}a (i)) and mass (Figure \ref{fig:time_and_mass}a (ii)) of the last $> 4 \times 10^{20}$ kg impactor for the 92\% of impact histories that do not experience later ocean vaporization. The median timing of a life-starting impact is 4.34 Ga with 95\% uncertainty between 3.89 and 4.46 Ga. The large uncertainty mostly stems from the stochastic, Poisson nature of asteroid bombardment. The probability distribution for the mass of the life-starting impact peaks at $4 \times 10^{20}$ kg and decreases with increasing mass because of the size-frequency distribution of asteroids (Figure \ref{fig:sfd_and_flux}a). There is only a 10\% chance that the impactor is Moneta-sized (i.e., between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg).

@@

Our estimated timing for biopoiesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology \citep{Rosing_1999,Ohtomo_2014}. Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be generally compatible with our estimating timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga.

@@

It is temping to extrapolate our results beyond the early Earth to exoplanets, but we must do so with caution because planet's orbiting different stars may have bombardment histories unlike the Hadean. Planets in the habitable zone of a late M-type star during the stellar main-sequence phase were interior to the habitable zone for several hundred million years during the super luminous pre-main-sequence phase \citep{Luger_2015}. \citet{Lichtenberg_2022} use N-body simulations to show that, for planets hosted by late M-dwarfs, large asteroid impacts are not useful for prebiotic chemistry because big impacts most likely occur within 100 million years of planet formation during the stellar pre-main-sequence when the planet is outside of the habitable zone. The pre-main-sequence phase is not an issue for an impact-induced origin of life on planets orbiting sun-like stars because the phase only lasts $\sim 3$ Myrs, while large asteroid impacts occur for 100s Myrs \citep{Lichtenberg_2022}. Therefore, our result that a life-starting impact was likely on the early Earth (92\% chance in an optimistic case) might be best extrapolated to habitable zone exoplanets orbiting sun-like stars.

@@

For either the optimistic or pessimistic cases, we find that the most likely timing for an origin of life impactor is $\sim 4.35$ Ga with a 95\% uncertainty spanning the entire Hadean eon (approximately 4.45 to 3.9 Ga). The large uncertainty is caused by the intrinsic stochastic nature of impacts. These results suggest that the \citet{Benner_2020} proposed timing of the origin of life, at $4.36 \pm 0.1$ Ga, is too narrow. Furthermore, the mass of the life starting impactor is most likely (69\% to 90\% probability) smaller than the Moneta impactor (Figure \ref{fig:time_and_mass}) proposed by \citet{Benner_2020}, because the size frequency distribution of asteroids prefers more frequent smaller impactors.

@@

We assume that the size-frequency distribution of objects that struck the early Earth is similar to the main-belt asteroids. A problem with this approach is that the main-belt asteroids only contain objects up to $\sim$ 1000 km diameter, yet the early Earth is expected to have experience impacts larger than this \citep{Marchi_2014}. Therefore, we must extrapolate the size-frequency distribution above $\sim$ 1000 km to larger objects. Throughout the main text we use the same extrapolation as \citet{Marchi_2014} (the black dashed line in Figure \ref{fig:sfd_and_flux}), who extends the size-frequency distribution with a log-linear slope of $d (\ln S_0)/d (\ln m) = - 0.415$.

To test the sensitivity of our results to this chosen extrapolation, we re-did our Monte-Carlo analysis using a log-linear slope of $d (\ln S_0)/d (\ln m) = - 1$ (the red dashed line in Figure \ref{fig:sfd_and_flux}) for impactors bigger than $\sim$ 1000 km diameter. Figures \ref{fig:probabilities_of_impacts_extrapolation_sens} and \ref{fig:timing_extrapolation_sensitivity} show how our results change when adopting this alternative extrapolation. Overall, our results are qualitatively unchanged for the different extrapolation slope ($d (\ln S_0)/d (\ln m)$).

@@

Recently, \citet{Citron_2022} performed smoothed-particle hydrodynamics impact simulations over a range of impact angles, masses and velocities to estimate the impact properties that can vaporize an ocean. Figure \ref{fig:energy_for_ocean_vap}a shows their simulation results for the change in the atmosphere's and ocean's internal energy ($\Delta \mathrm{IE}_\mathrm{atmo}$) as a function of impact kinetic energy ($\mathrm{E}_\mathrm{imp}$), along with log-linear extrapolations for each impact angle. Figure \ref{fig:energy_for_ocean_vap} gives the same information, but the y-axis is instead the fraction of the impactor's energy that heats the atmosphere and ocean over the course of their simulations. For the most probable incident impact angle of $45^\circ$ about $\sim 10\%$ of the impactors kinetic energy heats the atmosphere ($f_\mathrm{E,vap} = 0.1$), so we nominally adopt this value in the main text to determine which collisions vaporize the ocean.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/energy_for_ocean_vap.pdf}
  \caption{\citet{Citron_2022} smoothed-particle hydrodynamics simulations that show how much impactor kinetic energy heats the atmosphere and ocean. Different colors indicated various incident impact angles. (a) shows the simulated change in internal energy of the atmosphere and ocean as a function of the impactor energy, with log-linear extrapolations for each impact angle. The dotted horizontal line at $5 \times 10^{27}$ J indicates the energy needed to vaporize an ocean \citep{Sleep_1989}. (b) contains the same information as (a), except the y-axis is the fraction of impactor energy that heats the atmosphere and ocean.}
  \label{fig:energy_for_ocean_vap}
\end{figure}

@@

Figure \ref{fig:timing_sensitivity} shows the timing of the last impact to make conditions favorable for biopoiesis that does not experience subsequent ocean vaporization for $f_\mathrm{E,vap}$ values of 0.025, 0.1 (nominal value), and 0.25. Like in the main text, we consider the \citet{Wogan_2023} optimistic (Figure \ref{fig:timing_sensitivity}a) and pessimistic (Figure \ref{fig:timing_sensitivity}b) minimum impactor masses to produce significant prebiotic molecules. In Figure \ref{fig:timing_sensitivity}a, the timing of the last life-starting impact is relatively insensitive to $f_\mathrm{E,vap}$. In the \citet{Wogan_2023} pessimistic case (Figure \ref{fig:timing_sensitivity}b), an origin of life is preferred later in the Hadean for larger values of $f_\mathrm{E,vap}$.

>>>>

Here, we use the \citet{Wogan_2023} results, along with Monte-Carlo simulations of Earth's impact history, to make an alternative estimate for when life most likely emerged during an RNA-first process. Our calculations account for the possibility of planet sterilization by impacts that vaporize the ocean \citep{Sleep_1989}. We assume that a life-starting impact is one that produces significant prebiotic nitriles and is not subsequently followed by an ocean-vaporizing impact that destroys the biosphere. By considering the fraction of stochastic impact realizations that do not have a life-starting impact, we also estimate the probability of life beginning if Earth history was rerun.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and, in the main text, log-log extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. In Appendix \ref{sec:append_sfd} we also consider the red dashed extrapolation to $\gtrsim 10^{21}$ kg impacts. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

@@

\begin{equation}
  N_0 = A_\oplus s_0 n_m = A_\oplus s_0 10^{a e^{b t} + c}
  \label{eq:num_earth_impacts}
\end{equation}
Equation \eqref{eq:num_earth_impacts} also accounts for Earth's surface area ($A_\oplus$), making $N_0$ the number of impacts on Earth since age $t$ that would cause a crater $> 1$ km on the moon. As discussed previously, \citet{Morbidelli_2018} use crater scaling relations to show that a 1 km lunar crater corresponds to a 50 m object with mass $m_0 = 1.64 \times 10^8$ kg.

@@

\begin{equation}
  \overline{N}_{ij} = N(m_{j-\frac{1}{2}},t_{i-\frac{1}{2}},t_{i+\frac{1}{2}}) - N(m_{j+\frac{1}{2}},t_{i-\frac{1}{2}},t_{i+\frac{1}{2}})
\end{equation}
Next, we sample a Poisson distribution for each $\overline{N}_{ij}$, giving a stochastic number of impacts in each mass and time grid cell which constitutes an impact history. By modeling impacts as a Poisson process, we are assuming that each impact does not change the probability of the next one (i.e., impacts are independent). Performing this sampling thousands of times captures many of the possible impact histories. We only keep impact histories that accrete a total mass of $2 \times 10^{22}$ to $6 \times 10^{22}$ kg because this is approximately the mass implied by the HSEs in Earth's mantle, assuming that HSEs were delivered by late accretion. This range of masses is based on \citet{Day_2015} who used mantle HSEs to suggest that the Earth was impacted by $\sim 3 \times 10^{22}$ to $4.8 \times 10^{22}$ kg of material, but we choose a wider range of masses because HSEs could have been lost to Earth's core or to space during impacts \citep{Marchi_2018}.

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Appendix Figure \ref{fig:velocity_distribution} shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close-approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10 km$). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would cause biopoiesis depends on the required minimum impact mass to produce significant origin-of-life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} uses photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} finds that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.

@@

As illustrated in Figure \ref{fig:example_impact_histories}, a life-starting impact does not occur in every simulated impact history. In our impact-driven model for the origin of life, biopoiesis requires (1) an impact of sufficient mass to make prebiotic nitriles and (2) the lack of a later ocean-vaporizing collision that destroys the biosphere without rekindling it. Figure \ref{fig:probabilities_of_impacts} shows the probability of both conditions occurring as a function of the minimum impact mass that produces prebiotic molecules. For example, consider the \citet{Wogan_2023} optimistic minimum mass of $4 \times 10^{20}$ kg indicated in Figure \ref{fig:probabilities_of_impacts}. All simulated impact histories have at least one impact bigger than $4 \times 10^{20}$ kg (blue line in Figure \ref{fig:probabilities_of_impacts} is 1.0 at the given mass). However, the probability of the origin of life in this scenario is 92\%, because 8\% of the time the last $4 \times 10^{20}$ kg collision is followed by a smaller ocean-vaporizing impact that perhaps sterilizes the planet (orange dashed line in Figure \ref{fig:probabilities_of_impacts} with 0.92 at $4 \times 10^{20}$ kg). An origin of life is less probable if we consider the \citet{Wogan_2023} pessimistic minimum mass ($> 5 \times 10^{21}$ kg) to produce significant origin-of-life nitriles. A $> 5 \times 10^{21}$ kg impact without later ocean vaporization occurs 41\% of the time. A life-starting impact was therefore relatively likely on the early Earth, and was strongly favored (92\% probability) under optimistic assumptions.

@@

To estimate the most likely timing for the origin of life, we consider both the optimistic and pessimistic criteria of \citet{Wogan_2023}. Figure \ref{fig:time_and_mass}a is the optimistic case, showing the timing (Figure \ref{fig:time_and_mass}a (i)) and mass (Figure \ref{fig:time_and_mass}a (ii)) of the last $> 4 \times 10^{20}$ kg impactor for the 92\% of impact histories that do not experience later ocean vaporization. The median timing of a life-starting impact is 4.34 Ga with 95\% uncertainty between 3.89 and 4.46 Ga. The large uncertainty mostly stems from the stochastic, Poisson nature of asteroid bombardment. The probability distribution for the mass of the life-starting impact peaks at $4 \times 10^{20}$ kg and decreases with increasing mass because of the size-frequency distribution of asteroids (Figure \ref{fig:sfd_and_flux}a). There is only a 10\% chance that the impactor is Moneta-sized (i.e., between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg).

@@

Our estimated timing for biopoiesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology \citep{Rosing_1999,Ohtomo_2014}. Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be generally compatible with our estimated timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga.

@@

It is temping to extrapolate our results beyond the early Earth to exoplanets, but we must do so with caution because planets orbiting different stars may have bombardment histories unlike the Hadean. Planets in the habitable zone of a late M-type star during the stellar main-sequence phase were interior to the habitable zone for several hundred million years during the super luminous pre-main-sequence phase \citep{Luger_2015}. \citet{Lichtenberg_2022} use N-body simulations to show that, for planets hosted by late M-dwarfs, large asteroid impacts are not useful for prebiotic chemistry because big impacts most likely occur within 100 million years of planet formation during the stellar pre-main-sequence when the planet is outside of the habitable zone. The pre-main-sequence phase is not an issue for an impact-induced origin of life on planets orbiting sun-like stars because the phase only lasts $\sim 3$ Myrs, while large asteroid impacts occur for 100s Myrs \citep{Lichtenberg_2022}. Therefore, our result that a life-starting impact was likely on the early Earth (92\% chance in an optimistic case) might be best extrapolated to habitable zone exoplanets orbiting sun-like stars.

@@

For either the optimistic or pessimistic cases, we find that the most likely timing for an origin of life impactor is $\sim 4.35$ Ga with a 95\% uncertainty spanning the entire Hadean eon (approximately 4.45 to 3.9 Ga). The large uncertainty is caused by the intrinsic stochastic nature of impacts. These results suggest that the \citet{Benner_2020} proposed timing of the origin of life, at $4.36 \pm 0.1$ Ga, is too narrow. Furthermore, the mass of the life-starting impactor is most likely (69\% to 90\% probability) smaller than the Moneta impactor (Figure \ref{fig:time_and_mass}) proposed by \citet{Benner_2020}, because the size frequency distribution of asteroids prefers more frequent smaller impactors.

@@

We assume that the size-frequency distribution of objects that struck the early Earth is similar to the main-belt asteroids. A problem with this approach is that the main-belt asteroids only contain objects up to $\sim$ 1000 km diameter, yet the early Earth is expected to have experience impacts larger than this \citep{Marchi_2014}. Therefore, we must extrapolate the size-frequency distribution above $\sim$ 1000 km to larger objects. Throughout the main text we use the same extrapolation as \citet{Marchi_2014} (the black dashed line in Figure \ref{fig:sfd_and_flux}), who extends the size-frequency distribution with a log-log slope of $d (\ln S_0)/d (\ln m) = - 0.415$.

To test the sensitivity of our results to this chosen extrapolation, we re-did our Monte-Carlo analysis using a log-log slope of $d (\ln S_0)/d (\ln m) = - 1$ (the red dashed line in Figure \ref{fig:sfd_and_flux}) for impactors bigger than $\sim$ 1000 km diameter. Figures \ref{fig:probabilities_of_impacts_extrapolation_sens} and \ref{fig:timing_extrapolation_sensitivity} show how our results change when adopting this alternative extrapolation. Overall, our results are qualitatively unchanged for the different extrapolation slope ($d (\ln S_0)/d (\ln m)$).

@@

Recently, \citet{Citron_2022} performed smoothed-particle hydrodynamics impact simulations over a range of impact angles, masses and velocities to estimate the impact properties that can vaporize an ocean. Figure \ref{fig:energy_for_ocean_vap}a shows their simulation results for the change in the atmosphere's and ocean's internal energy ($\Delta \mathrm{IE}_\mathrm{atmo}$) as a function of impact kinetic energy ($\mathrm{E}_\mathrm{imp}$), along with log-log extrapolations for each impact angle. Figure \ref{fig:energy_for_ocean_vap} gives the same information, but the y-axis is instead the fraction of the impactor's energy that heats the atmosphere and ocean over the course of their simulations. For the most probable incident impact angle of $45^\circ$ about $\sim 10\%$ of the impactors kinetic energy heats the atmosphere ($f_\mathrm{E,vap} = 0.1$), so we nominally adopt this value in the main text to determine which collisions vaporize the ocean.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/energy_for_ocean_vap.pdf}
  \caption{\citet{Citron_2022} smoothed-particle hydrodynamics simulations that show how much impactor kinetic energy heats the atmosphere and ocean. Different colors indicated various incident impact angles. (a) shows the simulated change in internal energy of the atmosphere and ocean as a function of the impactor energy, with linear log-log extrapolations for each impact angle. The dotted horizontal line at $5 \times 10^{27}$ J indicates the energy needed to vaporize an ocean \citep{Sleep_1989}. (b) contains the same information as (a), except the y-axis is the fraction of impactor energy that heats the atmosphere and ocean.}
  \label{fig:energy_for_ocean_vap}
\end{figure}

@@

Figure \ref{fig:timing_sensitivity} shows the timing of the last impact to make conditions favorable for biopoiesis that does not experience subsequent ocean vaporization for $f_\mathrm{E,vap}$ values of 0.025, 0.1 (nominal value), and 0.25. As in the main text, we consider the \citet{Wogan_2023} optimistic (Figure \ref{fig:timing_sensitivity}a) and pessimistic (Figure \ref{fig:timing_sensitivity}b) minimum impactor masses to produce significant prebiotic molecules. In Figure \ref{fig:timing_sensitivity}a, the timing of the last life-starting impact is relatively insensitive to $f_\mathrm{E,vap}$. In the \citet{Wogan_2023} pessimistic case (Figure \ref{fig:timing_sensitivity}b), an origin of life is preferred later in the Hadean for larger values of $f_\mathrm{E,vap}$.
